Hunting experience and prey variability jointly shape individual foraging specialisation in a predator-prey videogame

### 2022-08-19 manuscript/Chapter3-manuscript.Rmd
update refs and introduction MS

The development of foraging strategies is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that expertise optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics when prey are scarce or highly unpredictable [@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that expertise may offset the costs/risks of switching foraging tactics as individuals gain experience and information on their prey or their environment, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013]. In this case, switching between foraging tactics would be advantageous when resources are variable, as behavioral flexibility is often key for survival in novel environments [reviewed in @Snell-Rood2013]. *revoir cette phrase ça ressemble trop à l'autre. plutôt parler des processus cognitifs qui modulent cela*

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging. Importantly, we have little information on the ecological contexts - except for competition - that favors one strategy over the other. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specializing [@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. While identifying the relevant contexts mediating these fitness benefits is a daunting task, in predator-prey systems, some clues indicate that fluctuations in the predictability of prey (e.g. abundance, availability, behavior) encounters throughout a predator's lifetime may be a key factor mediating the benefits of specializing [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability hypothesis advances that when resources are predictable, individual specialists should have higher delivery rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should be advantaged when resources fluctuate, as fine adjustments to prey are key for a predator's success and survival.

>>>>

The development of foraging strategies is crucial for young predators to reach adulthood and survive [@Phillips.etal2017;@Heithaus.etal2018]. Theory predicts that foraging specialization may emerge via learning, limitations in memorizing multiple complex hunting skills, and expertise [@Tinker.etal2009;@Dukas2019]. @Dukas2019 defines expertise as the characteristics, skills, and knowledge allowing individuals with extensive experience to outperform novices on complex tasks. This body of work suggests that the development of expertise is an optimizing process that promotes foraging specialization. Empirical studies on human and non-human hunters show that expertise optimizes the efficiency (e.g. search and handling times, return rates) of their foraging tactics potentially via associative images or reliance on prey and environmental cues [@Edwards.Jackson1994;@Morse2000;@macdonaldCrossculturalComparisonLearning2007;@Reid.etal2010;@Wilson-Rankin2015]. Such optimization may thus reinforce the use of the same tactic (i.e. specialization) if its success is constant each time a prey is encountered. It may also be costly to attempt different hunting tactics by trial and error when prey are scarce or highly unpredictable [@Dukas1998;@Estes.etal2003;@Mery.Burns2010]. An alternative mechanism is that expertise may offset the costs/risks of switching foraging tactics as individuals gain experience and information on their prey or their environment, leading to increased individual foraging flexibility [@Ishii.Shimada2010;@Mery.Burns2010;@Snell-Rood2013]. To develop proper responses to fluctuating resources, individuals would need to sample their environment broadly during an extensive period of time, followed by higher performance at later stages of development [reviewed in @Snell-Rood2013].

There is currently a lack of consensus on the fitness advantages of specialized vs flexible foraging. Importantly, we have little information on the ecological contexts - except for competition - that favors one strategy over the other. This is reflected in the literature showing contrasting results in the links between specialization and fitness. For instance, some studies report increasing benefits of specializing [@Patrick.Weimerskirch2014a;@vandenBosch.etal2019], some report that flexible foraging has greater benefits [@Paull.etal2012;@Manlick.etal2021], and others find equal benefits depending on timescales [@Woo.etal2008;@Potier.etal2015]. While quantifying the fitness consequences of specialization is a daunting task, in predator-prey systems, some clues indicate that fluctuations in the predictability of prey (e.g. abundance, availability, behavior) encounters throughout a predator's lifetime may be a key factor [@Weimerskirch2007;@Woo.etal2008;@Chang.etal2017;@Phillips.etal2017;@Courbin.etal2018]. The resource-predictability hypothesis advances that when resources are predictable, individual specialists should have higher delivery rates by reducing the energy and time required to search for and handle prey. In contrast, individual generalists should be advantaged when resources fluctuate, as fine adjustments to resources are key for a predator's success and survival [@Karkarey.etal2017;@Holm.etal2019;@Santoro.etal2019].
